\section[An introduction to the red thread for this thesis]{An introduction to the red thread for this thesis}
\label{section_read_thread} 

\epigraph{{An effective \textcolor{red}{red}} thread stop examiners from seeing `red'}{Anon (1964--on)}

A `red thread'~\footnotemark 
includes the core message of the thesis that's intended to keep the reader on track throughout the thesis. Here it's being used primarily to help me keep my work on message and enable me to discard topics and material that are too far removed from the red thread. This chapter is not written for publication, however some of the contents may be incorporated into the overall thesis.

\footnotetext{The term red thread is used in Scandinavian schools according to \href{https://writingcooperative.com/red-thread-and-other-writing-guidelines-4910b0d0d395}{‘Red Thread’ and Other Writing Guidelines}. An example of applying a red thread to a thesis is provided in \href{https://patthomson.net/2018/04/02/thesis-knowhow-how-the-contribution-can-create-coherence/}{thesis knowhow – “the contribution” can create coherence}. Tips on applying a red thread to improve a thesis are provided by a highly experienced PhD examiner in \href{https://threadreaderapp.com/thread/1192531751949737986.html}{Thread}. Others use a different colour, e.g. a \href{https://phdinahundredsteps.com/2015/05/26/spinning-the-golden-thread-that-can-sew-your-phd-together/}{a golden thread that can sew a PhD together}, and it may simply be a thread as follows: \href{https://www.thephdproofreaders.com/structure-a-phd/how-to-find-the-thread-that-runs-through-your-phd-thesis/}{How to find the thread that runs through your PhD thesis}.}

A key maxim for me is to: \emph{Always connect software engineering, mobile analytics, with the aim of improving reliability of mobile apps for end users.}



\clearpage
\section{Red Thread for the introduction chapter}~\label{red-thread-introduction}
Main chapter: \secref{chapter-introduction}.

Mobile apps are ubiquitous and used by billions. These apps, %10 words
the software tools, and their ecosystem are all imperfect therefore failures will occur and some will adversely affect the user experience and have knock-on effects. %25 words

This research uses mobile analytics to help developers of real world apps understand reliability failures, address the causes, and deploy new releases with improved reliability. %25 words
Source, value, and impact are also considered of doing so, as are real-world constraints. %15 words


\section{Red Thread for the Literature Review}
Main chapter: \secref{chapter-related-work}.


\section[Research methodology]{Research methodology}
Main chapter: \secref{chapter-methodology}. 

Empirical studies are highly appropriate and therefore used. They include complimentary primary and secondary research methods. Common threads facilitate comparisons and generalising aspects of the research. %26 words


\clearpage
\input{empirical-studies/empirical-studies-red-thread}
%\href{section-empirical-studies-red-thread}{\nameref{section-empirical-studies-red-thread}} includes a table of the role the researcher played in each case study together with other pertinent details.

\subsection{Observations on the case studies}
The research includes several aspects seldom studied owing to the challenges of obtaining the case studies, for instance from behind the curtain in commercial projects with over a million users, and from sources again seldom visible to researchers \emph{i.e.} of mobile analytics for real-world mobile apps.


\clearpage 
\subsection{When can identified issues be addressed?}
To address issues they need to be identified and a practical improvement determined. The practical improvement may be an outright fix, or a workaround, or an amelioration, etc. The questions for the development team include where and when the improvements can be applied to improve the reliability for the end users of their app.

Some issues can be ameliorated externally to the app, for instance though changes to servers that support API requests from apps, however many need changes to the app in order to be effective.

For apps released as compiled binary files (the vast majority in iOS and Android app stores), improvements are generally made to a subsequent release than the one(s) where the failures have been occurring\footnotemark. And these subsequent releases need to be installed by the user population and used similarly in order to determine whether the improvements were worthwhile. Some users keep older releases and others stop using the app so 100\% rollouts of subsequent releases are not practical. (For apps that only ever had one release then 100\% rollout is \emph{de-facto}.)

\footnotetext{
  There are various ways that a binary \emph{can} be patched or updated \textit{in-situ}. Indeed various security attacks are performed using this approach, as, for instance,~\citep{poeplau2014_execute_this_unsafe_android} discussed. Around April 2013 Google changed their policy for Google Play to forbid Android apps from modifying, replacing or updating it's own APK binary code outside Google Play's update mechanism (the policy is still available on the Internet Archive Wayback Machine~\href{https://web.archive.org/web/20130426113437/https://play.google.com/about/developer-content-policy.html}{Google Play Developer Program Policies}). In 2019, Google provided developers with an API that developers can embed into their Android app in order to prompt users to update the app using Google Play~\citep{android_in_app_updates}. 
}



\newthought{Norms per app category}
Video conferencing apps may have a higher mean crash rate than apps in other categories. 

\noindent
\rule{\textwidth}{0.4pt}

\clearpage
\section{Where's my current focus?}
It's in several areas:
\begin{itemize}
    \item Create a presentation to sum up the research.
    \item Provide coherence and consistency to the empirical studies. Learn about methodologies applied to empirical studies.
    \item (Ongoing) Some additional development of the possible framework in Section 0.6 and Figure 0.3
    \item (Ongoing) On the related work chapter, adding various references and starting to develop these into coherent discussions of these works, the gaps in research, etc.
    \item (Done) To revise the red-thread for the case studies and create three distinct tables as per my discussion with Marian and then flesh out and organise the case studies.
\end{itemize}

%%% I learned how to add sub-titles to sections thanks to
% https://tex.stackexchange.com/questions/69550/how-can-i-add-a-subtitle-to-a-section-title
% If I want to get fancy then https://tex.stackexchange.com/questions/417429/how-to-make-subheadings 
