\chapter{Case Studies}
\section{Overview of Case Studies}
The case studies provide %three %MUST_DO revise the introduction once the number and characteristics of the case studies is settled.
distinct perspectives of applying Mobile Analytics to Android apps. Android apps were selected to enable the use of Google Play Console reports and analytics, including Android Vitals; nonetheless one of the case studies chose to also include analytics in their iOS app so this will also be covered briefly.

Early research explored ways developers of opensource Android apps use local logging, a complementary and oft used approach intended to help developers learn more about how their app behaves locally at run-time. 

\begin{table}[th]
    \centering
    \begin{tabular}{l|l|r|l}
      Case Study &Role of Researcher &Apps \\
      \hline
       \href{https://play.google.com/store/apps/dev?id=9116215767541857492&hl=en_GB}{Kiwix}  &Embedded &18 \\
       \href{https://play.google.com/store/apps/developer?id=Catrobat&hl=en_GB}{Catrobat} &Guide &6 \\
       \href{https://play.google.com/store/apps/dev?id=7665838187257770408}{Greentech Apps} &\textbf{TBD} &10 \\
       \href{https://play.google.com/store/apps/developer?id=Moonpig.com&hl=en_GB}{Moonpig.com} &Observed &1 \\
       \href{https://play.google.com/store/apps/details?id=boundless.moodgym&hl=en_GB}{Moodspace app} &Interviewed &1 \\
       \href{https://play.google.com/store/apps/details?id=com.localhalo.app&hl=en_GB}{Local Halo app} &Observed &1 \\
    \end{tabular}
    \caption{Project teams and Commercial apps in the case studies}
    \label{tab:case_studies}
\end{table}

For the Kiwix case study, the researcher has been an intrinsic long-term member of the diffuse project team, able to work directly with the code-base and collaborate directly with the developers and ancillary members of the project team. 

For the Catrobat case study, the researcher advised and assisted the project team to apply mobile analytics to their larger, older, and less reliable app: \emph{Pocket Code}. The researcher helped lead a one-day hackathon and otherwise interacted through a bug reporting tool, JIRA, discussions and using shared documents. Pocket Code also included a crash-reporting library which allowed cross-tool comparisons of reports, analytics and data. During the research, the reporting platform for the crash-reporting was migrated to a newer service which provided further insights and comparisons across and between the various mobile analytics tools.

For the commercial app teams (Moonpig, Moodspace, and LocalHalo), the researcher corresponded with one of the development team for each of the commercial apps and received either direct access to their analytics tools (LocalHalo), or was provided with snapshots (Moonpig and Moodspace). Permission was granted by their respective organisations for their contributions to be used for research purposes.
\newpage

\input{case-studies/kiwix-team-case-study}
\newpage

\input{case-studies/catrobat-case-study}
\newpage

\input{case-studies/gtaf-case-study}
\newpage

\input{case-studies/developers-experiences-of-mobile-analytics}

\section{Research in logging practices}

\begin{itemize}
    \item Our research.
    \item The opensource tools we created to facilitate the testing and analysis of logging by Android developers.
    \item Explorations in methods to improve the effectiveness of logging and the analysis of log messages generated by apps.
\end{itemize}




\section{Summary of Case Studies}
The case studies includes a useful range of Android apps developed by independent teams using a variety of programming languages, mindsets, objectives, and constraints. In each team they learned to actively focus on stability metrics as reported in various technology-facing analytics tools, and the developers continue to see the merit of doing so on an ongoing basis.