\chapter{Future work}
This PhD provides a snapshot of research I have done in recent years. This work fits within a larger body of research with similar overall objectives, to improve both the software and how that software is developed. This chapter introduces various research I am actively involved in which builds on the work described in my thesis.

\section{Evaluating Google Play Console Reports}
With discovering the many and various flaws and quirks in Google's analytics products I decided a set of experiments would help increase the trust and transparency of Android Vitals and Google Play Console. Aspects of these experiments may also be highly relevant to evaluate other analytics tools, products, and services. Some of the experiments could trigger defensive mechanisms used to 'protect' large-scale online services such as Google Play. For instance App Stores, including Google Play, are known to take protective and corrective activities related to fake reviews. Some experiments might want to assess the effectiveness of protective defenses, others may aim to avoid triggering them. 

Google is known to take punitive actions against developers whom it deems to be related to undesirable behaviours. %MUST-DO add references to examples. 
As these actions can include a lifetime ban for the account deemed to have transgressed, and for other accounts it considers are related, the effects of their actions can be far reaching and remove all revenues for an organisation's Android apps where the onus is on the people in that organisation being able to convince Google of the merits of reinstatement often without knowing what the reasons were, what the evidence was, and with no other practical way of restoring their account.

The Google Engineering team have been defensive and guarded in terms of suggestions to help evaluate the behaviours of their system. SHOULD-DO add some examples. 

A cautious set of basic tests to help establish characteristics of analytics from no (zero) to tens of users is available online at 
\url{https://joedocs.com/julianharty/evaluating-gpc-reports}

\textbf{Placeholders to extend on this topic}: Holding ourselves and each other to account. Humility, openness, comparisons with financial auditing of businesses. \emph{c.f.} Sentry.io's work, opensourcing of client libraries, etc. 

\begin{itemize}
    \item Trust is earned, not claimed.
    \item All it takes for evil to flourish quote?
\end{itemize}

\section{Enhancing quality vs. enhancing UX}

\section{Adding in-app analytics to PocketCode}

\section{Logging in mobile apps}
My research in mobile analytics overlaps and closely aligns with research on logging - both are used by software development teams to learn about how their software behaves. I am part of a group of distributed, international researchers investigating aspects of how and why developers add logging to their mobile applications. One of our current research areas investigates the use of Google Firebase Analytics for logging. \emph{Ad-hoc} research notes are available at \url{https://joedocs.com/julianharty/apm-logging-research}. 
