\section{Publications}
My various public contributions follow in this section.

\subsection{Peer-reviewed papers}
\begin{enumerate}
    \item 	Giovanna Avellis, Julian Harty, Yijun Yu. \emph{``Towards mobile twin peaks for app development"}. In MOBILESoft '17 Proceedings of the 4th International Conference on Mobile Software Engineering and Systems. 2017. \href{https://doi.org/10.1109/MOBILESoft.2017.10}{DOI 10.1109/MOBILESoft.2017.10}~\cite{avellis_harty_yu_towards_mobile_twin_peaks}
    
    \textbf{My contribution:} I co-authored the work and presented the work at the conference.
    \item Julian Harty. \emph{``Google Play Console: Insightful Development using Android Vitals and Pre-Launch Reports"}. In MOBILESoft '19 Proceedings of the 6th International Conference on Mobile Software Engineering and Systems. 2019. \href{https://doi.org/10.1109/MOBILESoft.2019.00019}{DOI: 10.1109/MOBILESoft.2019.00019} ~\cite{harty_google_play_console_insightful_development_using_android_vitals_and_pre_launch_reports}
    
    \textbf{My contribution:} I wrote and presented the paper and a poster at the conference.
    
    \item Julian Harty, Matthias Müller. \emph{``Better Android Apps using Android Vitals"}.  WAMA 2019: Proceedings of the 3rd ACM SIGSOFT International Workshop on App Market Analytics August 2019 Pages 26–32 \href{https://doi.org/10.1145/3340496.3342761}{DOI 10.1145/3340496.3342761}~\cite{harty_better_android_apps_using_android_vitals}
    
    \textbf{My contribution:} I wrote most of the paper and did most of the research. Joseph Reeve developed the Vitals Scraper software with my input. Matthias Müller contributed the section on Catrobat.
    
    \item HUCAPP 2020 \emph{``Stuck In Limbo With Magical Solutions: The Testers’ Lived Experiences of Tools and Automation"}
    
    \textbf{My contribution:} I helped inspire the research and contributed to revising and improving the paper.
    
    \item ICST 2020 Doctoral Symposium \emph{``How Can Software Testing be Improved by Analytics to Deliver Better Apps?"}
    
    \textbf{My contribution:} I wrote the paper to present my research from a testing perspective.
    
    \item TAIC-PART 2020 \emph{``Fast Abstract: Data Dynamics for Testing Systems"}
    
    \textbf{My contribution:} I wrote this paper, quickly to honour the concept of a fast abstract, on some of the many testing challenges related to data and data dynamics.
    
    \item TAIC-PART 2020 Joint paper \emph{``Test Tools: an illusion of usability?"}
    
    \textbf{My contribution:} I contributed some of the writing and ideas and helped write and format the paper for submission and publication. 
    
    \item MOBILESoft 2020 SRC \emph{``Improving App Quality Despite Flawed Mobile Analytics"}
    
    \textbf{My contribution:} I wrote the paper for the Student Research Competition.
    
    \item ICGSE 2020 \emph{``Designing Engineering Onboarding for 60+ Nationalities"}
    
    \textbf{My contribution:} I wrote the paper `shepherded' by an experienced author in the field of Global Software Engineering research.
    
\end{enumerate}
% How to create numbered lists https://www.overleaf.com/learn/latex/Lists

\subsection{Private reports}
\begin{enumerate}

    \setcounter{enumi}{9}
    \item Julian Harty. \emph{``Report on flaws discovered in Google Play Console and Android Vitals"}.
    
    \textbf{My contribution:} I wrote this report to assist the Google Engineering Team responsible for Google Play Console and Android Vitals at their request.

\end{enumerate}

\subsection{Books}
\begin{enumerate}
    \setcounter{enumi}{10}
    \item Julian Harty, Antoine Aymer. \emph{``The Mobile Analytics Playbook: A Practical Guide to Better Testing"}. Hewlett Packard Enterprise, 2015."
    ISBN: 978-0-9970694-0-2 (PDF) 978-0-9970694-1-9 (Print)
    
    \textbf{My contribution:} I wrote most of the book and did much of the background research.
    
    \item \emph{``The Mobile Developer's Guide to the Galaxy"} series of books, from \nth{6} to \nth{18} Editions. A widely distributed and read book that introduces the many and various challenges a developer of mobile apps needs to consider. The series is a collaborative project with multiple authors and editors where the book is published in print and electronically. The content is released under a creative commons license. 
    
    \textbf{My contributions:} I variously co-edited the book, wrote and revised various chapters both solely and with other authors. My main chapters include: Testing, Mobile Analytics, and Collecting and Understanding User Feedback. The current edition is freely available online at~\url{https://www.open-xchange.com/resources/mobile-developers-guide-to-the-galaxy/}
\end{enumerate}

\subsection{Software}
Please see the ~\hyperlink{software-contributions-chapter}{\emph{\nameref{software-contributions-chapter}}} chapter in the appendices for details. 
\clearpage
